A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2022-01-07 paper/elsarticle/elsarticle-template.tex
paper

@@

\begin{figure}
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projection to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table). (a) Original image, which we wish to project to the scene plane. (b) Projecting the original image to the scene plane from the viewpoint of the projector directed to the target location (bottom left, black with up vector red) introduces projective distortions if the projector is not fronto-parallel with the scene plane (as is the case here), with orientation and scale varying as a function of target location.}
    \label{fig:proj}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projection to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to a given target location in the scene plane, correcting for projective distortions using our approach. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach with respect to the given target location. (b) Projecting the warped image to the scene plane from the viewpoint of the projector directed to the target location (bottom left, black with up vector red; same as in Figure~\ref{fig:proj}(b)) has the effect of projecting the \textit{original} image from the viewpoint of the \textit{virtual} projector (center, gray with up vector red). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

@@

One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface \cite{Hartley2004}. The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane. Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector, which they use to carry out a rotational correction in projecting to a wall. Still another way to compute a homography is as a function of a plane and a pair of pinhole cameras; the effect produced by the homography is one of projecting an image to the plane from the viewpoint of the one camera, and acquiring the projected image from the viewpoint of the other. Such a homography is said to be `induced' by a plane \cite{Hartley2004}. We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, yet additionally taking into account both desired in-plane orientation and metric scale.

@@

@@

\begin{figure}[ht]
    \centering
    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.15cm]{images/detections_.png}}}
    \qquad
    \subfloat[\centering Corresponding 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.5cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the `downwards-facing' camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Circle pattern center points and chessboard corners detected automatically (overlain for illustration). (b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$. These are obtained using bundle adjustment by minimizing the cumulative reprojection error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ gives the projection of the point~$\mathbf{X}_{i,j} \in \mathbb{R}^3$ to the image plane of the camera underlying the $j^{\text{th}}$~calibration image, subject to the maximum likelihood estimates $\hat{\mathtt{K}}, (\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$. The function~$d_{\hat{\omega}}$, in turn, is a function that first (i) corrects $\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'$ for lens distortions in accordance with the maximum likelihood estimate of the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}, and then (ii) computes a distance in pixels with respect to two resulting corrected positions. More generally, the projection~$\mathbf{x}' \in \mathbb{R}^2$ of a point~$\mathbf{X} \in \mathbb{R}^3$ to the image plane of a camera given calibration matrix~$\mathtt{K}$ and pose~$(\mathtt{R}, \mathbf{t})$ is given by
\begin{equation}
(\mathbf{x}'^\top, 1)^\top \sim \mathtt{K}(\mathtt{R}\mathbf{X} + \mathbf{t}) \in \mathbb{P}^2,
\end{equation}
where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses $\mathbf{X}$ in the coordinate frame of the camera. Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera then concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the `downwards-facing' camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).

@@

Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the `downwards-facing' camera. Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location). This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. Our physical `downwards-facing' camera---in terms of whose rotation relative to the scene plane we wish to orient our augmentations---is almost certain to not face downwards precisely. Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation. It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

@@

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figures~\ref{fig:warp} and \ref{fig:final_proj}) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out---again intuitively---with respect to the ground plane.

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=4.75cm]{images/system3_.pdf}}}  %3.6
    \quad
    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=4.75cm]{images/meas2_.pdf}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation, to the upper-left corner of our floorspace---an image warped using our approach. The digital angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge. Corners of the full projected image extent projected to the floor in light green to provide context. Note that blur has a minimal impact on the quality of the augmentation, and hence on the accuracy of the measurements.}
    \label{fig:eval}
\end{figure}

>>>>

\begin{figure}[t]
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.65cm]{images/projim.png}}}
    \quad
    \subfloat[\centering Projection of original image to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.7cm]{images/proj_projim.png}}}
    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table). (a) Original image, which we wish to project to the scene plane. (b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector is not fronto-parallel with the scene plane (as is the case here), with orientation and scale varying as a function of target location.}
    \label{fig:proj}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.65cm]{images/projim_warped_scaled.png}}}
    \quad
    \subfloat[\centering Projection of our warped image to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.7cm]{images/proj_projim_waped.png}}}
    \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized using our approach. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach. (b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

@@

@@

One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface \cite{Hartley2004}. The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane. Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity (i.e., down) vector, which they use to carry out a rotational correction in projecting to a wall. Still another way to compute a homography is as a function of a plane and a pair of pinhole cameras; the effect produced by the homography is one of projecting an image to the plane from the viewpoint of the one camera, and acquiring the projected image from the viewpoint of the other. Such a homography is said to be `induced' by a plane \cite{Hartley2004}. We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, yet additionally taking into account both desired in-plane orientation and metric scale.

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.15cm]{images/detections_.png}}}
    \quad
    \subfloat[\centering Recovered 3D positions corresponding to projector calibration image for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.7cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the `downwards-facing' camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Circle pattern center points and chessboard corners detected automatically (overlain for illustration). (b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

@@

A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j), j \in \{ 1, \dots, n_\text{im} \}$. These are obtained using bundle adjustment by minimizing the cumulative reprojection error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ gives the projection of the point~$\mathbf{X}_{i,j} \in \mathbb{R}^3$ to the image plane of the camera underlying the $j^{\text{th}}$~calibration image, subject to the maximum likelihood estimates $\hat{\mathtt{K}}, (\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$. The function~$d_{\hat{\omega}}$, in turn, is a function that first (i) corrects $\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'$ for lens distortions in accordance with the maximum likelihood estimate of the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}, and then (ii) computes a distance in pixels with respect to two resulting corrected positions. More generally, the projection~$\mathbf{x}' \in \mathbb{R}^2$ of a point~$\mathbf{X} \in \mathbb{R}^3$ to the image plane of a camera given calibration matrix~$\mathtt{K}$ and pose~$(\mathtt{R}, \mathbf{t})$ is given by
\begin{equation}
(\mathbf{x}'^\top, 1)^\top \sim \mathtt{K}(\mathtt{R}\mathbf{X} + \mathbf{t}) \in \mathbb{P}^2,
\end{equation}
where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses $\mathbf{X}$ in the coordinate frame of the camera. Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.

Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera then concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the `downwards-facing' camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$, for each projector calibration image).

@@

Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the `downwards-facing' camera. Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location). It is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. Our physical `downwards-facing' camera---placed to face downwards towards the scene plane, and in terms of whose rotation relative to the normal vector of the scene plane we wish to orient our augmentations---is almost certain to not face downwards towards the scene plane precisely. Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation relating the two.} %It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely. The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

@@

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t})$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figures~\ref{fig:warp} and \ref{fig:final_proj}) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out---again intuitively---with respect to the ground plane.

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup consisting of projector, steerable mirror system, `downwards-facing' camera (i.e., camera placed to face downards towards the scene plane).]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
    \quad
    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=5cm]{images/meas2_.pdf}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation, to the upper-left corner of our floorspace---an image warped using our approach. The digital angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge. Corners of the full projected image extent projected to the floor in light green to provide context. Note that blur has a minimal impact on the quality of the augmentation, and hence on the accuracy of the measurements.}
    \label{fig:eval}
\end{figure}
